---
title: Syllabus
slug: syllabus
citeproc: false
bibliography: ../static/bib/references.bib
csl: ../static/bib/chicago-syllabus-no-bib.csl
---

```{r load-libraries, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

[^courseinfo]: {-}
  `r blogdown::shortcode("courseinfo")`
 
# Course Objectives

The course has two related goals: (1) Provide a theoretical understanding of the scientific process and the role of cumulative science, and (2) introduce you to a set of modern tools for conducting science in this framework. Specifically, you'll be able to:

- Understand the basic principles of experimental design and the cumulative scientific process 
- Use R/RStudio and the tidyverse framework (Wickham, 2017) for conducting basic data analysis 
- Present visualizations of data using Grammar of Graphics principles (Wickham, 2010)
- Implement reproducible workflows
- Conduct a basic meta-analysis

In addition, the course will introduce you to a range of other tools ([Github](https://github.com/), [Rpubs](http://rpubs.com/), [Open Science Framework](https://osf.io/), web experimental frameworks, metafor) that are useful in psychological research. 

# Course Overview

The course is organized into four units. The __first unit__ will introduce the philosophy of cumulative science and the main themes of the course. The __second unit__ will focus on tools for conducting a single experiment. In this unit, we'll learn tools for analyzing and visualizing data within the tidyverse (Wickham, 2017) framework. We'll also learn tools for creating a reproducible workflow in experimental research. In the __third unit__, we'll introduce the Null Hypothesis Testing framework and the concept of replication. As a class, we'll conduct an online replication of a single experiment. Finally, the __fourth unit__ will focus on a quantitative approach to synthesizing the results of many experiments: meta-analysis. The basic priciples of meta-analysis will be introduced, and students will conduct an original meta-analysis as a final project.

# Course materials

The readings in this class are free and will be provided and the course website. There is no book for the course.

The course will make heavy use of R and RStudio, free at
http://cran.r-project.org/ and http://www.rstudio.com/.

# Requirements
This course assumes no prior programming experience, but does require a willingness to learn. Prior familiarity with any programming language (e.g. R, Python, Matlab), however, will be helpful in the course.

# Policies

## Participation

Completing the readings and attending class are mandatory. Lecture slides will be posted after class each day, but reading these slides is only a partial substitute for class attendance. The material we will be covering requires hands-on involvement, and you will get the most out this class if you actively engage in the material, particularly in lab. 

## Late Work
Late work policy for the problem sets and lab reports:

  - next day: lose 30% of total possible points
  - later than next day: lose all points
  
Late work policy for the project: 10% off for each day late.

Regrade requests must be made within one week of when the assignment is returned, and must be submitted in writing. These will be honored if points were tallied incorrectly, or if you feel your answer is correct but it was marked wrong. No regrade will be made to alter the number of points deducted for a mistake.

## Cheating and Plagiarism: 

Cheating and plagiarism are defined in the CMU Student Handbook, and include (1) submitting work that is not your own for papers, assignments, or exams; (2) copying ideas, words, or graphics from a published or unpublished source without appropriate citation; (3) submitting or using falsified data; and (4) submitting the same work for credit in two courses without prior consent of both instructors. Any student who is found cheating or plagiarizing on any work for this course will receive a failing grade for that work. Further action may be taken if necessary, including a report to the dean.


# Assignments and grades

<span class="newthought">You can find descriptions</span> for all the assignments on the [assignments page](/assignments/).

```{r assignments-grades, include=FALSE}
assignments <- tribble(
  ~Points,  ~Assignment,
  16      , "Participation",
  8 * 20  , "Lab Assignments (8 × 20)",
  48     , "Midterm",
  96     , "Final Meta-analysis Project [Proposal = 16; Presentation = 30; Writeup = 50]",
) %>%
  mutate(Percent = Points / sum(Points),
         Percent = scales::percent(Percent)) 

grading <- tribble(
  ~Grade, ~Range,
  "A",  "90–100%",
  "B",  "80–89%",
   "C",  "70-79%", 
  "D",  "60-69%"
)
```

```{r show-assignments-table, echo=FALSE, results="asis"}
assignments_total <- tibble(Assignment = "Total",
                            Points = sum(assignments$Points),
                            Percent = NA)

bind_rows(assignments, assignments_total) %>% 
  select(Assignment, Percent, Points) %>% 
  pandoc.table(justify = "lcc", missing = "")
```

\

```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lc", 
               col.names = c("Grade", "Range"))
```

