---
title: "ggplot example: gapminder <br>"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/sta199-logo-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
 library(gapminder)
library(knitr)
#knitr::opts_chunk$set(fig.height = 2, fig.width = 3, dpi = 300) 


```

class: center, middle


```{css, echo=FALSE}
/* custom.css */
.left-code {
  color: #777;
  width: 38%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 60%;
  float: right;
  padding-left: 1%;
}
.plot-callout {
  height: 225px;
  width: 450px;
  bottom: 5%;
  right: 5%;
  position: absolute;
  padding: 0px;
  z-index: 100;
}
.plot-callout img {
  width: 100%;
  border: 4px solid #23373B;
}
```




## Data: `gapminder` dataset
```{r, echo = F}
gapminder <- gapminder %>%
  janitor::clean_names()
kable(head(gapminder, 10), "html")
      
```




---

```{r raw, eval = F}
ggplot()
```

--

.right-plot[
```{r raw, ref.label="thedata", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

```{r thedata, eval = F}
ggplot(data = gapminder)
```


--

.right-plot[
```{r thedata, ref.label="thedata", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---


```{r mappin, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp))
```

--

.right-plot[
```{r mappin, ref.label="mappin", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

Let's add a geom (note the `+`).

--

```{r geom, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp)) +
  geom_point()
```

--

.right-plot[
```{r, ref.label="geom", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]


---

Let's try a different geom: a smoothed line.

--

```{r smooth, eval = F, message = F, warning = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp)) +
  geom_smooth()
```

--

.right-plot[
```{r, ref.label="smooth", echo=FALSE, fig.dim=c(8, 6), out.width="100%", message = F, warning = F}
```
]

---

Maybe we want a linear line.

--

```{r linear, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp)) +
  geom_smooth(method = "lm")
```

--

.right-plot[
```{r, ref.label="linear", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

It might be nice to see the raw data WITH the line. We can combine geoms!

--

```{r line, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp)) +
  geom_point() +
  geom_smooth(method = "lm")
```

--

.right-plot[
```{r, ref.label="line", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]


---
Those points are too big. Let's make them smaller. Let's also change the shape. (why don't these go in aesthetics?).

--

```{r size, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp)) +
  geom_point(size = .8, shape = 1) +
  geom_smooth(method = "lm")
```

--

.right-plot[
```{r, ref.label="size", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]



---

Let's switch the variables back. How could we add information about year?

--

Through the color aesthetic...

--

```{r thiscolor, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp,
                                       color = year)) +
  geom_point(size = .8, shape = 1) +
  geom_smooth(method = "lm")
```

--

.right-plot[
```{r, ref.label="thiscolor", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

The data are all squished on the x axis. We can fix that by changing the scale on the x-axis to the be the log of x. 

--

```{r log, eval = F}
ggplot(data = gapminder, mapping = aes(x = gdp_percap, 
                                       y = life_exp,
                                       color = year)) +
  geom_point(size = .8, shape = 1) +
  geom_smooth(method = "lm") +
  scale_x_log10() 
```

--

.right-plot[
```{r, ref.label="log", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---


---

What if wanted to look at just data for the United States? 

--

Filter!

--

```{r us, eval = F}
gapminder %>%
  filter(country == "United States") %>%
  ggplot(mapping = aes(x = gdp_percap, y = life_exp)) +
  geom_point(size = .8) +
  geom_smooth(method = "lm") 
```

--

.right-plot[
```{r, ref.label="us", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

What if wanted to label the points with years?

--

`geom_label()`

--

```{r years, eval = F}
gapminder %>%
  filter(country == "United States") %>%
  ggplot(mapping = aes(x = gdp_percap, y = life_exp, label = year)) +
  geom_label() +
  geom_smooth(method = "lm") 
```

--
.right-plot[
```{r, ref.label="years", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]

---

Let's do a few things to make our plot more readable: Add clearer axis labels and a title. 

--


```{r readable, eval = F}
gapminder %>%
  filter(country == "United States") %>%
  ggplot(mapping = aes(x = gdp_percap, 
                       y = life_exp,
                       label = year)) +
  geom_label() +
  geom_smooth(method = "lm")  +
  xlab("GDP (per capita)") +
  ylab("Life expectancy (years)") +
  ggtitle("Life expectancy vs GDP in US")
```

--

.right-plot[
```{r, ref.label="readable", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]
---

We can also change the *theme*. I like `theme_classic`.

--


```{r theme, eval=FALSE}
gapminder %>%
  filter(country == "United States") %>%
  ggplot(mapping = aes(x = gdp_percap, 
                       y = life_exp, 
                       label = year)) +
  geom_label() +
  geom_smooth(method = "lm")  +
  xlab("GDP (per capita)") +
  ylab("Life expectancy (years)") +
  ggtitle("Life expectancy vs GDP in US") +
  theme_classic(base_size = 16)
```

--

.right-plot[
```{r ref.label="theme", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```
]
---




So far we've just been working with the raw data in tidy format. We can also *summarize* our data before we plot it. 

--

What if we wanted to know which country had the highest life expectancy? The second highest?

--

```{r country, eval = F}
life_exp_by_country <- gapminder %>%
  group_by(country) %>%
  summarize(mean_life_exp = mean(life_exp))

ggplot(life_exp_by_country, mapping = aes(x = country,
                                          y = mean_life_exp)) +
  geom_bar(stat = "identity")
```

---


```{r, ref.label="country", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```



Yowza! That's a disaster!

---

Here's a prettier version:

```{r country2, eval = F}
life_exp_by_country <- gapminder %>%
  group_by(continent, country) %>%
  summarize(mean_life_exp = mean(life_exp),
            mean_pop = mean(pop)) %>%
  filter(mean_pop > 10000000)


ggplot(life_exp_by_country, aes(x = fct_reorder(country, mean_life_exp) ,
                                          y = mean_life_exp,
                                          fill = continent)) +
  geom_bar(stat = "identity") +
  ylab("Mean Life expectancy (years)") +
  xlab("Country") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```

---

```{r, ref.label="country2", echo=FALSE, fig.dim=c(8, 6), out.width="100%"}
```


