---
title: "Calculating Confidence Intervals in R <br>"
subtitle: "Modern Research Methods"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/sta199-logo-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
 library(gapminder)
library(knitr)
library(janitor)
knitr::opts_chunk$set(fig.height = 2, fig.width = 3, dpi = 300) 
```


```{css, echo=FALSE}
/* custom.css */
.left-code {
  color: #777;
  width: 38%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 60%;
  float: right;
  padding-left: 1%;
}
.plot-callout {
  height: 225px;
  width: 450px;
  bottom: 5%;
  right: 5%;
  position: absolute;
  padding: 0px;
  z-index: 100;
}
.plot-callout img {
  width: 100%;
  border: 4px solid #23373B;
}
```


We're going to calculate a confidence intervals for the means on accuracy reported in Zettersten and Lupyan (2020), Experiment 1B

Let's start by loading the data.
```{r, cache = T, message = F, warning = F}
DATA_PATH<- "https://osf.io/a4dzb/download"
 
zl_data <- read_csv(DATA_PATH)

zl_clean <- zl_data %>%
     clean_names() %>%
     select(experiment, subject, age, condition, block_num, is_right)

zl_exp1a <- zl_clean %>%
     filter(experiment == "1A")
```

```{r, echo = F}
zl_exp1a %>%
  slice(1:5) %>%
  kable(format = "html")
```


---

We start by getting by-subject by-condition means
```{r}
ms_by_overall<- zl_exp1a %>%
  group_by(subject, condition) %>%
  summarize(prop_right = sum(is_right)/n()) %>%
  ungroup() 
```

--

```{r, echo = F}
ms_by_overall %>%
  slice(1:5) %>%
  kable(format = "html")
```

---

Next, let's calculate a **point estimate** for the mean in each condition.
```{r}
means_by_condition  <- ms_by_overall %>%
  group_by(condition) %>%
  summarize(mean = mean(prop_right))
```

--

```{r, echo = F}
kable(means_by_condition, format = "html")
```

---
Plot the point estimates with geom_point. 
```{r}
ggplot(means_by_condition, aes(x = condition, y = mean)) +
  geom_point(size = 2) + 
  ylim(.5, 1) +
  theme_classic()
```


---

Next let's calculate a confidence interval around our estimate. 

To start we need the sample size in each condition.
```{r}
sample_size <- ms_by_overall %>%
  group_by(condition) %>%
  summarize(n = n())
```

--

```{r, echo = F}
kable(sample_size,  format = "html")
```

---

Now, let's calculate the the CI
```{r}
means_by_condition_with_ci <- ms_by_overall %>%
  group_by(condition) %>%
  summarize(mean = mean(prop_right),
            sd = sd(prop_right)) %>%
  mutate(ci_range_95 =  1.96 * (sd/sqrt(25)), #<<
         ci_lower = mean - ci_range_95,
         ci_upper = mean + ci_range_95)
```

--

```{r, echo = F}
kable(means_by_condition_with_ci, format = "html")
```

---

Plotting the confidence intervals 
```{r}
ggplot(means_by_condition_with_ci, aes(x = condition, y = mean)) +
  geom_point(size = 2) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + #<<
  ylim(.5, 1) +
  theme_classic()
```

---

There's actually a single geom that plots both points and ranges: geom_pointrange.  
```{r}
ggplot(means_by_condition_with_ci, aes(x = condition, y = mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + #<<
  ylim(.5, 1) +
  theme_classic()
```


---

There's one small complexity that I've glossed over. 


--

Because we don't actually know the SD for the population distribution we have to estimate from a distribution called the t-distribution. 
```{r}
means_by_condition_with_ci_t <- ms_by_overall %>%
  group_by(condition) %>%
  summarize(mean = mean(prop_right),
            sd = sd(prop_right)) %>%
  mutate(ci_range_95 =  qt(1 - (0.05 / 2), 25 - 1) * (sd/sqrt(25)), #<<
         ci_lower = mean - ci_range_95,
         ci_upper = mean + ci_range_95)
```

--

```{r, echo = F}
kable(means_by_condition_with_ci_t, format = "html")
```

---

Point estimates with ranges calculated from the t-distribution. 

```{r}
ggplot(means_by_condition_with_ci_t, aes(x = condition, y = mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + #<<
  ylim(.5, 1) +
  theme_classic()
```
