---
title: "Calculating and Plotting Effect Sizes <br>"
subtitle: "Modern Research Methods"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/sta199-logo-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(janitor)
library(compute.es)
knitr::opts_chunk$set(fig.height = 2, fig.width = 3, dpi = 300) 
theme_set(theme_classic(base_size = 6)) 
```


```{css, echo=FALSE}
/* custom.css */
.left-code {
  color: #777;
  width: 38%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 60%;
  float: right;
  padding-left: 1%;
}
.plot-callout {
  height: 225px;
  width: 450px;
  bottom: 5%;
  right: 5%;
  position: absolute;
  padding: 0px;
  z-index: 100;
}
.plot-callout img {
  width: 100%;
  border: 4px solid #23373B;
}
```


Let's calculate an effect size for Zettersten and Lupyan, 1A and our "replication".

Start by loading the data.

--

```{r, cache = T, message = F, warning = F}
DATA_PATH <- "https://osf.io/a4dzb/download"
 
zl_data <- read_csv(DATA_PATH)

zl_clean <- zl_data %>%
     clean_names() %>%
     select(experiment, subject, age, condition, block_num, is_right)

zl_exp1a <- zl_clean %>%
     filter(experiment == "1A")

ms_by_overall<- zl_exp1a %>%
  group_by(subject, condition) %>%
  summarize(prop_right = sum(is_right)/n())  %>%
  mutate(experiment = "original_ZL2020") %>%
  ungroup()
```

---

Our data
```{r, message = F}
ms_by_overall_replication <- read_csv("mrm_replication_data.csv") %>%
  rename(subject = subid) %>% #<<
  mutate(subject = as.character(subject))
```

--


Combine our replication data and the original.
```{r}

all_data <- bind_rows(ms_by_overall, ms_by_overall_replication) %>% #<<
  select(experiment, subject, everything()) #<< 
```


---
The data:
```{r, echo = F}
all_data %>%
  slice(1:5) %>%
  kable(format = "html")
```

---
Let's do a sanity check and look at the number of subjects in each condition/ experiment

--

```{r}
all_data %>%
  group_by(experiment, condition) %>%
  summarize(n = n())
```

--


Or, equivalently:
```{r}
all_data %>%
  count(experiment, condition) 
```

---

Let's plot the two distributions now. 


```{r, eval = F}
ggplot(all_data, aes(x = prop_right, fill = condition)) +
  geom_density(alpha = .5) +
  facet_wrap(~experiment) 
```

--

```{r,  echo = F, fig.height = 1.5}
ggplot(all_data, aes(x = prop_right, fill = condition)) +
  geom_density(alpha = .5) +
  facet_wrap(~experiment) 
```

---

Let's get summary statistics for each experiment/condition

```{r}
dist_summary <- all_data %>%
  group_by(experiment, condition) %>%
  summarize(mean_prop_right = mean(prop_right),
            sd = sd(prop_right),
            n = n())
```

--

```{r, echo = F}
kable(dist_summary,
      format = "html")
```

---

Let's use these new data frame to add the means to our plot. 

```{r,  fig.height = 1.5, eval = F}
ggplot(all_data, aes(x = prop_right, fill = condition)) +
  geom_density(alpha = .5) +
  geom_vline(data = dist_summary, aes(xintercept = mean_prop_right), linetype = 2) + #<<
  facet_wrap(~experiment)
```

--

```{r,  fig.height = 1.5, echo = F}
ggplot(all_data, aes(x = prop_right, fill = condition)) +
  geom_density(alpha = .5) +
  geom_vline(data = dist_summary, aes(xintercept = mean_prop_right), linetype = 2) + #<<
  facet_wrap(~experiment)
```

---

## We're now ready to calculate the effect size in each experiment.

--
+ Functions in the `compute.es` package allow us to easily calculate effect size measures.
--

+ To calculate cohen's d (the effect size for the difference between two means), we use the `mes()` function.
--

+ The syntax for `mes()` is: 

```{r, eval = F}
mes(m.1,  # mean in first group
    m.2, # mean in second group
    sd.1, # sd of first group
    sd.2, # sd of second group
    n.1, # sample size of first group
    n.2) # sample size of second group
```

---

To use the `mes` function, were going to need to "pull" out values from our `dist_summary` data frame. 
```{r, echo = F}
dist_summary %>%
  kable(format = "html")
```
--

```{r}
dist_summary %>% 
  filter(experiment == "MRM_replication_of_ZL2020", condition == "high") %>% 
  pull(mean_prop_right)

```

---



```{r}
replication_effect_size <-
  mes(dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "high") %>% pull(mean_prop_right), # m.1
    dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "low") %>% pull(mean_prop_right),# m.2
    dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "low") %>% pull(sd), #sd.1
    dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "high") %>% pull(sd), #sd.2
    dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "low") %>% pull(n), #n.1
    dist_summary %>% filter(experiment == "MRM_replication_of_ZL2020", condition == "high") %>% pull(n), #n.1
    verbose = F) %>%
  mutate(experiment = "MRM_replication_of_ZL2020")
```

--

```{r}
replication_effect_size
```

---

```{r}
original_effect_size <-
  mes(dist_summary %>% filter(experiment == "original_ZL2020", condition == "high") %>% pull(mean_prop_right),
    dist_summary %>% filter(experiment == "original_ZL2020", condition == "low") %>% pull(mean_prop_right),
    dist_summary %>% filter(experiment == "original_ZL2020", condition == "low") %>% pull(sd),
    dist_summary %>% filter(experiment == "original_ZL2020", condition == "high") %>% pull(sd),
    dist_summary %>% filter(experiment == "original_ZL2020", condition == "low") %>% pull(n),
    dist_summary %>% filter(experiment == "original_ZL2020", condition == "high") %>% pull(n),
    verbose = F) %>%
  mutate(experiment = "original_ZL2020")
```

--

```{r}
original_effect_size
```


---

Let's combine the two data frames

```{r}
both_es <- bind_rows(original_effect_size, replication_effect_size)
```

--

```{r}
both_es
```

---

And, tidy them: 
```{r}
tidy_es <- both_es %>%
  select(experiment, d, l.d, u.d) %>%
  rename(ci_lower = l.d,
         ci_upper = u.d)
```

--

```{r}
tidy_es
```


---

And, plot them!
```{r,  fig.height = 1.5, eval = F}
ggplot(tidy_es, aes(y = d, x= experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylim(0,2) +
  ggtitle("Effect size in original and replication")
```

--

```{r,  fig.height = 1.5, echo = F}
ggplot(tidy_es, aes(y = d, x= experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylim(0,2) +

  ggtitle("Effect size in original and replication")
```


---

```{r,  fig.height = 1.5, eval = F}
ggplot(tidy_es, aes(y = d, x = experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Effect size in original and replication") +
  ylim(0,2) +
  coord_flip()
```

--

```{r,  fig.height = 1.5, echo = F}
ggplot(tidy_es, aes(y = d, x = experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Effect size in original and replication") +
  ylim(0,2) +
  coord_flip()
```


---

## Interpretation:

+ The effect size is numerically bigger in the original experiment (compared to the replication)
+ But, the confidence intervals are overlapping, meaning that they have similiar estimates for plausible values for the population effect size.

```{r,  fig.height = 1.5, echo = F}
ggplot(tidy_es, aes(y = d, x = experiment)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Effect size in original and replication") +
  ylim(0,2) +
  coord_flip()
```
