---
title: Syllabus
slug: syllabus
citeproc: true
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-fullnote-bibliography-no-bib.csl
output:
  blogdown::html_page:
    toc: true
    fig_caption: false
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
```

`r blogdown::shortcode("courseinfo")`

# Course Objectives

The course has two related goals: (1) Provide a theoretical understanding of the scientific process and the role of cumulative science, and (2) introduce you to a set of modern tools for conducting science in this framework. Specifically, you'll be able to:

- Understand the role of replicability and reproducibility for cumulative psychological science
- Understand the principles of "the new statistics" (i.e., confidence intervals, effect sizes; Cumming, 2014)
- Conduct a basic meta-analysis
- Use R/RStudio and the tidyverse framework (Wickham, 2017) for conducting basic data analysis 
- Present visualizations of data using Grammar of Graphics principles (Wickham, 2010)

In addition, the course will introduce you to a range of other tools ([Github](https://github.com/), [Rpubs](http://rpubs.com/), [Open Science Framework](https://osf.io/), metafor) that are useful in psychological research. 

# Course Overview

The course is organized into four units. The __first unit__ will introduce the process of cumulative science and the main themes of the course. The __second unit__ will focus on tools for conducting a single experiment. In this unit, we'll learn tools for analyzing and visualizing data within the tidyverse (Wickham, 2017) framework. We'll also learn tools for creating a reproducible workflow in experimental research. In the __third unit__, we'll introduce the concept of replication and related statistical concepts. Finally, the __fourth unit__ will focus on a quantitative approach to synthesizing the results of many experiments: meta-analysis. The basic principles of meta-analysis will be introduced, and students will conduct an original meta-analysis as a final project.

# Course materials

The readings in this class are free and will be provided on the course website. There is no book for the course.

The course will make heavy use of [R](http://cran.r-project.org/) and [RStudio](http://www.rstudio.com/). R is the programming software and RStudio is the dashboard (IDE).  Both of these are open source and freely available. You'll need to install both of these on your computer. If you're having trouble with this, please reach out to either the TA or myself. This software is already installed on most CMU lab computers.

# Requirements
This course assumes no prior programming experience, but does require a willingness to learn and be challenged. Prior familiarity with any programming language (e.g. R, Python, Matlab), however, will be helpful in the course.

# Policies

## Classroom Expectations related to COVID-19

The goal is to create an environment for learning that will also keep everyone as safe as possible. Therefore, in order to attend class meetings, I expect all students to abide by all behaviors indicated in [A Tartan’s Responsibility](https://www.cmu.edu/coronavirus/students/tartans-responsibility.html), including any timely updates based on the current conditions. In particular, I expect students to wear a facial covering throughout class when there is a requirement in effect on campus (the requirement is in effect at the start of the Fall 2021 semester). 

## Participation

Learning a new "language" (in this case, R) and the other skills in this course require hands-on involvement, and you will get the most out this class if you actively engage in the material.  Students are therefore expected to complete readings and participate in class.  Lecture slides will be posted after class each day, but reading these slides is only a partial substitute for class attendance. 

##  Laptops 
You will need to bring a laptop to class for Monday quizzes and Friday labs. If this is a problem, please let me know.  To facilitate learning for everyone, the expectation is that laptops will not be not be used in lecture outside of quizzes. 
 
## Late Work
Late work policy for the assignments, midterm, and final:

  - next day: lose 25% of total possible points
  - later than next day: lose all points
  
Your lowest assignment score will be dropped. Late work policy for the final project: 10% off for each day late.

## Take care of yourself

The pandemic has introduced  a lot of stress and uncertainty for everyone over the last year and a half. Make sure to take care of yourself this semester.  Move regularly, eat well, and reach out to your support system or me if you need to. We can all benefit from support in times of stress, and that is especially true this semester. 

As a student, you may experience a range of challenges that can interfere with learning, such as strained relationships, increased anxiety, substance use, feeling down, difficulty concentrating and/or lack of motivation. These mental health concerns or stressful events (like a global pandemic!) may diminish your academic performance and/or reduce your ability to participate in daily activities. It is important to take care of yourself by seeking help, just as you would for any physical health issue CMU services are available, and treatment does work. You can learn more about confidential mental health services available on campus at: http://www.cmu.edu/counseling/. Support is always available (24/7) from Counseling and Psychological Services: 412-268-2922.

If you are worried about affording food or feeling insecure about food, there are resources on campus who can help. Email the [CMU Food Pantry Coordinator](cmu-pantry@andrew.cmu.edu) to schedule an appointment or call 412-268-8704 (SLICE office).

## Accommodations for Students with Disabilities

If you have a disability and have an accommodations letter from the Disability Resources office, I encourage you to discuss your accommodations and needs with me as early in the semester as possible. I will work with you to ensure that accommodations are provided as appropriate. If you suspect that you may have a disability and would benefit from accommodations but are not yet registered with the Office of Disability Resources, I encourage you to contact them at access@andrew.cmu.edu.

## Respect in the Classroom
It is my intent that students from all diverse backgrounds and perspectives be well served by this course, that students’ learning needs be addressed both in and out of class, and that the diversity that students bring to this class be viewed as a resource, strength and benefit. It is my intent to present materials and activities that are respectful of diversity: gender, sexuality, disability, age, socioeconomic status, ethnicity, race, and culture. Your suggestions are encouraged and appreciated. Please let me know ways to improve the effectiveness of the course for you personally or for other students or student groups. In addition, if any of our class meetings conflict with your religious events, please let me know so that we can make arrangements for you.


## Cheating and Plagiarism 

Cheating and plagiarism are defined in the CMU Student Handbook, and include (1) submitting work that is not your own for assignments or exams; (2) copying ideas, words, or graphics from a published or unpublished source without appropriate citation; (3) submitting or using falsified data; and (4) submitting the same work for credit in two courses without prior consent of both instructors. Any student who is found cheating or plagiarizing on any work for this course will receive a failing grade for that work. Further action may be taken if necessary, including a report to the dean. 


# Assignments and grades

## Attendance and Participation
Attendance and participation (e.g., asking and answering questions) in lecture and lab are expected. To account for this, a portion of your final grade is based on your active participation. Your active participation in lecture and lab can earn you up to 20 points in your final grade.

<!---Nevertheless, I also recognize that students may need to miss class for a variety of reasons (religious observance, job interview, university-sanctioned event, or illness/quarantine). With that in mind, you can miss 3 class sessions (lectures or labs; e.g. 2 lectures and 1 lab) without any negative effect on your grade. When you must miss class, please notify me (at least 24 hours in advance except for illness/emergency), so that we can discuss alternative arrangements for catching up on class and associated work. If you encounter extenuating circumstances and must miss more than three classes, please come and discuss the issue with me; I would like to find a way to support you.--->

## Weekly quizzes
Most Mondays there will be a short 5 minute quiz at the beginning of class. These quizzes will cover the material from the previous two lectures (Monday and Wednesday lectures from the previous week). The goal of these quizzes is to give both you and I a sense of your understanding of the material.

##  Assignments 

There will be eight assignments throughout the semester. The goal of these assignments is to give you hands-on experience with the skills that are introduced in lecture and in lab. You are welcomed, and encouraged, to work with each other on the problems. But, you must turn in your own work.

## Midterm Exam
There will be a take-home midterm exam that will assess your conceptual understanding of the topics covered in the first part of the semester through short answer responses and R code. It will be similar in spirit to Assignments 1-5, but unlike the assignments, you will be expected to complete the exam completely on your own.

## Final project
The final project will be an original meta-analysis on a topic in cognitive, developmental or social psychology. You are encouraged to work on the project in groups of 3-5; however, each student must complete their own final write-up. Scope and topics of the meta-analysis will be decided in consultation with me.

There is no final exam. This project is your final exam.

## Assignment Grades

```{r assignments-grades, include=FALSE}
assignments <- tribble(
  ~Points,  ~Assignment,
  20      , "Participation",
  5*10, "Quizzes (5 x 10)",
  8 * 20  , "Lab Assignments (20 x 8)",
  50     , "Midterm",
  70, "Final Meta-analysis Project [Presentation = 35; Writeup = 35]",
) %>%
  mutate(Percent = Points / sum(Points),
         Percent = scales::percent(Percent)) 

grading <- tribble(
  ~Grade, ~Range,
  "A",  "90–100%",
  "B",  "80–89%",
   "C",  "70-79%", 
  "D",  "60-69%"
)
```

```{r show-assignments-table, echo=FALSE, results="asis"}
assignments_total <- tibble(Assignment = "Total",
                            Points = sum(assignments$Points),
                            Percent = NA)

bind_rows(assignments, assignments_total) %>% 
  select(Assignment, Percent, Points) %>% 
  pandoc.table(justify = "lrr", missing = "")
```

\

```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lc", 
               col.names = c("Grade", "Range"))
```

